---
title: "Penguin ANOVA"
author: "Tyler Cobian"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load Package: Tidyverse, car, onewaytests
```{r include=FALSE}
library(tidyverse)
library(car)
library(onewaytests)
penguins<- read_csv("penguins.csv")

```

Our question: Is there a significant difference between mean flipper length for flipper Chinstrap, Gentoo, Adelie penguins near Palmer Archipelago

Maybe I can use a one way ANOVA...
- Indipendant observations
- Equal variances (levene's test)
-Normality (histograms and qq plots)

```{r}
counts<- with(penguins, table(species, sex))

penguins2<- penguins %>% 
  mutate(
    common_name = 
      case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
    )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")
```

Exploratory graphs and a Levene's test

```{r}
hist<- ggplot(penguins2, aes(x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~ common_name, scale = "free")
hist

qq<- ggplot(penguins2, aes(sample = flipper_length))+
  geom_qq(aes(color = common_name))+
  facet_wrap(~common_name, scales = "free")
qq
#based on the Histogram and qqplot results data for each group appear normally distribusted. Also n is greater than 30, so means will be normally distributed by the CLT.
```
 
 
Evaluate th variances with a Levene's test for equal variances

```{r}
#If the sample variances is < 4x greater than the smallest sample, then these are "close enough" to consider the variances equal

variances<- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )
variances # The largest variance (33.1) is definitly not more than 4x greater than the smallest variance (15.2)

#Leven's test
#leveneTest function in the "car" package

levene_penguin<- leveneTest(flipper_length~common_name, data = penguins2)

#H0: no difference in variaces
#HA: There is a difference in variances

# There is no significant difference between species. 
```

One-way ANOVA for penguin flipper length by species (common_name)

```{r}
#H0: means across all groups are equal
#HA: At least two means differ significantly

penguin_aov<- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

#post-hoc test bu Tukey's HSD
#H0: There is not significant difference in pair wise means
#HA: At least two means differ significantly
penguin_ph<- TukeyHSD(penguin_aov)
penguin_ph

#aov.test function is the onewaytest packages allows for inline refrencing
```

Mean flipper leangth of female penguins differed significantly for all species studied (one-way ANOVA, F(df) = statistic, p<0.001, $\alpha$ = 0.05)


